---
title: "final189"
author: "Kalyani Cauwenberghs"
date: "3/7/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Part 1
#Part 2
Instructions:
For the final project, look back at all the analysis approaches you have used throughout the quarter. Consider HealthGen as the outcome, grouped into Excellent/Vgood, versus Good/Fair/Poor. (In the dataset dat.Rda, the HealthGen variable is assigned value 1 if its original value is Excellent/Vgood and is assigned value 0 otherwise.) Consider all other variables as potential predictors. Develop a comprehensive and reproducible analysis report, to explore the relationship between these variables and the outcome. Pay attention to (but not limited to) the following:

##1.)
Missing data: do not remove observations with any missing data from the start; after screening you might reduce to a smaller set of variables, therefore remove fewer observations at that point. Also you may consider removing variables with too much missing.
```{r}
load("dat (1).rda")
#remove columns with more than 310 missing
dat<-dat[,-which(colSums(is.na(dat)) > 310)]
#screening with univariate regression
screen<-function(var) {
  no_miss<-which(!is.na(var)& !is.na(dat$HealthGen))
  m<-glm(dat$HealthGen[no_miss]~var[no_miss], family = binomial())
  #true if p value < 0.05
  return(coef(summary(m))[2,4]<0.05)
}
var_indices<-which(colnames(dat) !="HealthGen")
#screen each column of dat excluding HealthGen
for (i in var_indices) {
  if(!screen(dat[,i])) {
    var_indices<-var_indices[which(var_indices!=i)]
  }
}
print("our final screened variables:")
colnames(dat)[var_indices]
```


##2.)
Include “Table 1”
```{r}
#do later.
```


##3.)
After univariate screening, building a multiple logistic regression model to predict the general health outcome of very good or excellent versus otherwise. State clearly your criteria at each step in the narrative.
```{r}

```


##4.)
Assess the predictability of the model by computing the (generalized) R-squared and the area under the ROC curve (AUC), as well as the cross-validated AUC.

##5.)
Use the variables that have passed the univariate screening, to build a classification tree. Describe clearly how you arrive at the final tree. Compute the error rate of your classification tree.

##6.)
Discuss any limitations in the analysis.

##Bonus
Explore random forest on the data above.

