---
title: "Week_6"
author: "Kalyani Cauwenberghs"
date: "2/14/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Part 1

##Part 2

##Part 3
Describe the distribution of the outcome variable, identify a main predictor that you’re interested in studying its effect on the outcome
```{r}
#read in dataset
data<-read.csv("hfi_cc_2019.csv")
data<-data[data$year=="2017",]
#predictor: political freedom
#outcome: economic freedom
#make sure both columns have no missing data
#sum(as.character(data$pf_score)=="-")==0
#sum(as.character(data$ef_score)=="-")==0

#distribution of outcome aka EF
hist(as.numeric(as.character(data$ef_score)))
```
Our outcome (EF score) is normal but left skewed, mean around 6.5, median around 7.
Our main predictor is PF score.
```{r}
hist(as.numeric(as.character(data$pf_score)))
```



Identify other variables (i.e. predictors, often called covariates) that might be related to the outcome or the main predictor, discuss these variables in the context of part 2 above of this assignment

political freedom variables can be our predictors.

pf_rol

pf_ss

pf_movement

pf_expression

pf_identity


Carry out univariate logistic regression of the outcome on each of the predictors including the main predictor, interpret the results in terms of odds ratio etc.

```{r}
#predictor: pf_rol, outcome: ef_score
pf_rol<-as.numeric(as.character(data$pf_rol))
ef_score<-as.numeric(as.character(data$ef_score))
plot(pf_rol,ef_score)
```





Fit a multiple logistic regression model by including more than one predictors, interpret the results in terms of conditional odds ratio etc.


