---
title: "Week_6"
author: "Kalyani Cauwenberghs"
date: "2/14/2020"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Part 1

##Part 2

##Part 3
#a.) Describe the distribution of the outcome variable, identify a main predictor that you’re interested in studying its effect on the outcome
```{r}
#read in dataset
data<-read.csv("hfi_cc_2019.csv")
data<-data[data$year=="2017",]
#predictor: political freedom
#outcome: economic freedom
#make sure both columns have no missing data
#sum(as.character(data$pf_score)=="-")==0
#sum(as.character(data$ef_score)=="-")==0

#distribution of outcome aka EF
hist(as.numeric(as.character(data$ef_score)))
```

Our outcome (EF score) is normal but left skewed, mean around 6.5, median around 7.
Our main predictor is PF score.

```{r}
hist(as.numeric(as.character(data$pf_score)))
```



#b.) Identify other variables (i.e. predictors, often called covariates) that might be related to the outcome or the main predictor discuss these variables in the context of part 2 above of this assignment.


- The variables that determine political freedom can be our predictors for economic freedom because they are not directly used in the calculation of economic freedom (pf_rol, pf_ss, pf_movement, pf_expression, pf_identity).
- TODO: discuss these variables in the context of part 2 above of this assignment


#c.) Carry out univariate logistic regression of the outcome on each of the predictors including the main predictor, interpret the results in terms of odds ratio etc.

```{r}
#make the columns into numerics
pf_rol<-as.numeric(as.character(data$pf_rol))
pf_ss<-as.numeric(as.character(data$pf_ss))
pf_movement<-as.numeric(as.character(data$pf_movement))
pf_expression<-as.numeric(as.character(data$pf_expression))
pf_identity<-as.numeric(as.character(data$pf_identity))
ef_score<-as.numeric(as.character(data$ef_score))
#predictor: pf_rol, outcome: ef_score
rol<-glm(ef_score~pf_rol)
rol$coefficients
#predictor: pf_ss, outcome: ef_score
ss<-glm(ef_score~pf_ss)
ss$coefficients
#predictor: pf_movement, outcome: ef_score
movement<-glm(ef_score~pf_movement)
movement$coefficients
#predictor: pf_expression, outcome: ef_score
expression<-glm(ef_score~pf_expression)
expression$coefficients
#predictor: pf_identity, outcome: ef_score
identity<-glm(ef_score~pf_identity)
identity$coefficients
```

TODO: interpret the results in terms of conditional odds ratio etc.



#d.) Fit a multiple logistic regression model by including more than one predictors, interpret the results in terms of conditional odds ratio etc.

```{r}
reg<-glm(ef_score~pf_rol+pf_ss+pf_movement+
           pf_expression+pf_identity)
```

Coefficients:

```{r}
reg$coefficients
```

TODO: interpret the results in terms of conditional odds ratio etc.


